\propertitle

### 2019-11-18 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-18T10:37:30+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

We illustrate possible uses of our proposed model in
\sect~\ref{sec:application}, by applying it to two concrete data sets:
\begin{enumerate*}[label=(\alph*)]
\item the activity of 65 neurons recorded for 20 min from a rat's Medial
  Entorhinal Cortex \citep{stensolaetal2012}, \item the activity of 159
  neurons recorded for 15 min from a macaque's Motor Cortex
  \citep{rostamietal2016_r2017}.
\end{enumerate*}
For each data set the model gives the most plausible frequency distribution
with which all levels of total activity of a larger network of 1\,000--10\,000
neurons appeared during the recording. These guessed frequency
distributions are distinctly different from the recorded ones for the
sampled neurons. For the first data set, for example, the frequency
distribution for the full network has two very distinct modes, both at low
activities, %($0.7\%$ and $2.6\%$ of simultaneously active neurons
% within $3\;\mathrm{ms}$,
see \fig~\ref{fig:plots_6mom}, whereas the frequency distribution for the
sample is monotonically decreasing with its maximum at zero activity. The
frequency distribution for the second data set doesn't have two modes but
still presents one prominent shoulder in its low-activity mode. Note that
these guessed features of the full network cannot be inferred by the
application of maximum-entropy \emph{to the sample alone}. Although we don't
formulate any hypotheses on why the larger network could have two distinct
low-activity regimes -- the purpose of the present work is only
methodological -- these results shows that the proposed method can lead to
the formulation or preliminary assessment of interesting hypotheses.

@@

We first calculate the distribution by using six moments. This number
already provides almost as much information as the full frequency
distribution of the sample, and at the same time illustrates the use of the
approach in questions of statistical sufficiency.
Figure~\ref{fig:plots_6mom} shows the resulting densities (that is,
distribution${}\times N$) for four example values of full-network sizes:
$N=n$, $N=1\,000$, $N=5\,000$, $N=10\,000$. The case $N=n$ corresponds to
applying the maximum-entropy method at the sample level; it can be observed
that with six moments it reproduces almost exactly the measured frequency
distribution. We discuss the case of unknown $N$ in \sect~\ref{sec:N}.
\begin{figure}[!p]
\centering
\includegraphics[width=\linewidth]{_scripts/paper_stensola_6mom_distributionsN_prior2.pdf}%
\\[3em]%
\includegraphics[width=\linewidth]{_scripts/paper_riehle_6mom_distributionsN_prior2.pdf}%
\caption{***}
\label{fig:plots_6mom}
\end{figure}
%maxent_solutions_different_moments_prior2_6moments.nb

@@

Consider the three sets of constraints $M_{2}$, $M_{4}$, $M_{5}$ consisting
in the first two, four, five moments. For a full network of size
$N=10\,000$, we obtain the following differences:
\begin{equation}
  \label{eq:diff_suff_N}
  \begin{split}
  \ydi(M_{4},M_{2}) &= 81\;\nat = 35\;\hart,
\\    \ydi(M_{5},M_{4}) &= 0.037\;\nat = 0.016\;\hart,
  \end{split}
\end{equation}
where the Hartley ($\hart$) denotes base-10 logarithms
\citep[\sect~C.4]{iso2009}. In words, the measured frequencies of the
sample activity are 35 orders of magnitude more probable under sufficiency
of the first four moments than under sufficiency of the first two only; but
they are about equally probable ($10^{0.016} = 1.04$) whether we consider
the first five moments to be sufficient or just the first four.

@@

The conclusions as regards \enquote{cooperativity} or \enquote{interaction}
that we reach by applying the maximum-entropy method to the full network,
as proposed here, are not only different from the application at the sample
level, but also visually clearer. The plot of \fig~\ref{fig:plots_2_4mom},
showing the two clearly different distributions constrained by two and four
moments, should be compared with the plot for the distributions obtained at
the sample level under the same constraints, shown in
\fig~\ref{fig:comparisons_marginals} as dashed lines with empty red
circles (\textcolor{myred}{\LARGE$\bm{\mathord{\circ}}$}) and empty blue
squares (\textcolor{mypurpleblue}{$\bm{\square}$}). The measured
frequencies are shown as filled black triangles
(\textcolor{black}{$\bm{\triangledown}$}). It is necessary to use a
logarithmic scale to see the differences, which appear mainly in the tail.
\begin{figure}[!b]
\centering
% \includegraphics[width=\linewidth]{_scripts/comparison_full.pdf}%
% \\[3em]%
\includegraphics[width=\linewidth]{_scripts/trunct_comparison_marginals.pdf}%
\caption{***}
\label{fig:comparisons_marginals}
\end{figure}
% comparison_ME_sample_full_levels.nb

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-18T13:35:04+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

We illustrate possible uses of our proposed model in
\sect~\ref{sec:application}, by applying it to two concrete data sets:
\begin{enumerate*}[label=(\alph*)]
\item the activity of 65 neurons recorded for 20 min from a rat's Medial
  Entorhinal Cortex \citep{stensolaetal2012}, \item the activity of 159
  neurons recorded for 15 min from a macaque's Motor Cortex
  \citep{rostamietal2016_r2017}.
\end{enumerate*}
For each data set the model gives the most plausible frequency distribution
with which all levels of total activity of a larger network of 1\,000--10\,000
neurons appeared during the recording. These guessed frequency
distributions are distinctly different from the recorded ones for the
sampled neurons. For the first data set, for example, the frequency
distribution for the full network has two very distinct modes, both at low
activities, %($0.7\%$ and $2.6\%$ of simultaneously active neurons
% within $3\;\mathrm{ms}$,
see \fig~\ref{fig:plots_5mom}, whereas the frequency distribution for the
sample is monotonically decreasing with its maximum at zero activity. The
frequency distribution for the second data set doesn't have two modes but
still presents one prominent shoulder in its low-activity mode. Note that
these guessed features of the full network cannot be inferred by the
application of maximum-entropy \emph{to the sample alone}. Although we don't
formulate any hypotheses on why the larger network could have two distinct
low-activity regimes -- the purpose of the present work is only
methodological -- these results shows that the proposed method can lead to
the formulation or preliminary assessment of interesting hypotheses.

@@

We first calculate the distribution by using five moments. This number
already provides almost as much information as the full frequency
distribution of the sample (see \sect~\ref{sec:marginalization}).
Figure~\ref{fig:plots_5mom} shows the resulting densities (that is,
distribution${}\times N$) for four example values of full-network sizes:
$N=n$, $N=1\,000$, $N=5\,000$, $N=10\,000$. The case $N=n$ corresponds to
applying the maximum-entropy method at the sample level; it can be observed
that with six moments it reproduces almost exactly the measured frequency
distribution. We discuss the case of unknown $N$ in \sect~\ref{sec:N}.
\begin{figure}[!p]
\centering
\includegraphics[width=\linewidth]{_scripts/newpaper_stensola_5mom_distributionsN_priorc.pdf}%
\\[3em]%
\includegraphics[width=\linewidth]{_scripts/newpaper_riehle_5mom_distributionsN_priorc.pdf}%
\caption{***}
\label{fig:plots_5mom}
\end{figure}
%maxent_solutions_different_moments_prior2_6moments.nb

@@

Consider the three sets of constraints $M_{2}$, $M_{4}$, $M_{5}$ consisting
of the first two, four, five moments. For a full network of size
$N=10\,000$, we obtain the following differences:
\begin{equation}
  \label{eq:diff_suff_N}
  \begin{split}
  \ydi(M_{4},M_{2}) &= 81\;\nat = 35\;\hart,
\\    \ydi(M_{5},M_{4}) &= 0.037\;\nat = 0.016\;\hart,
  \end{split}
\end{equation}
where the Hartley ($\hart$) denotes base-10 logarithms
\citep[\sect~C.4]{iso2009}. In words, the measured frequencies of the
sample activity are 35 orders of magnitude more probable under sufficiency
of the first four moments than under sufficiency of the first two only; but
they are about equally probable ($10^{0.016} = 1.04$) whether we consider
the first five moments to be sufficient or just the first four.

@@

The conclusions regarding \enquote{cooperativity} or \enquote{interaction}
that we reach by applying the maximum-entropy method to the full network,
as proposed here, are not only different from the application at the sample
level, but also visually clearer. The plot of \fig~\ref{fig:plots_2_4mom},
showing the two clearly different distributions constrained by two and four
moments, should be compared with the plot for the distributions obtained at
the sample level under the same constraints, shown in
\fig~\ref{fig:comparisons_marginals} as dashed lines with empty red
circles (\textcolor{myred}{\LARGE$\bm{\mathord{\circ}}$}) and empty blue
squares (\textcolor{mypurpleblue}{$\bm{\square}$}). The measured
frequencies are shown as filled black triangles
(\textcolor{black}{$\bm{\triangledown}$}). It is necessary to use a
logarithmic scale to see the differences, which appear mainly in the tail.
\begin{figure}[!b]
\centering
% \includegraphics[width=\linewidth]{_scripts/comparison_full.pdf}%
% \\[3em]%
\includegraphics[width=\linewidth]{_scripts/trunct_comparison_marginals.pdf}%
\caption{***}
\label{fig:comparisons_marginals}
\end{figure}
% comparison_ME_sample_full_levels.nb
